---
title: "Occupational Income"
author: "Tyler Brown"
date: "`r Sys.Date()`"
output: html_document
---

## Import Libraries
```{r}
library(tidyr)
library(dplyr)
```

## Read the Data
```{r}
url = 'https://raw.githubusercontent.com/AlphaCurse/Occupational-Income/main/OccupationalIncome.csv'
df = read.csv(url)
head(df)
```

## Tidy and Transform Data
I will change the name of each column and eliminate the row specifying qualitity of data.
```{r}
colnames(df)[1] = 'Occupation'
colnames(df)[2] = 'Workers'
colnames(df)[3] = 'Weekly_Income'
colnames(df)[4] = 'Male_Workers'
colnames(df)[5] = 'Male_Weekly_Income'
colnames(df)[6] = 'Female_Workers'
colnames(df)[7] = 'Female_Weekly_Income'
df = df %>%
  filter(!row_number() %in% c(1))
head(df)
```

I will change each column from chr to int while eliminating all commas in each column.
```{r}
df$Weekly_Income = as.integer(gsub(",","",df$Weekly_Income))
df$Male_Workers = as.integer(gsub(",","",df$Male_Workers))
df$Male_Weekly_Income = as.integer(gsub(",","",df$Male_Weekly_Income))
df$Female_Workers = as.integer(gsub(",","",df$Female_Workers))
df$Female_Weekly_Income = as.integer(gsub(",","",df$Female_Weekly_Income))
head(df)
```

## Analysis comparing the median weekly incomes for 535 different occupations to identify the difference in income between males and females.
To identify the difference in income between males and females, we can isolate the Occupation, Male_Weekly_Income and Female_Weekly_Income columns respectively. 
```{r}
new_df = df %>%
  select(Occupation, Male_Weekly_Income, Female_Weekly_Income)
head(new_df)
```

For consistency, any row with missing data will need to be removed.
```{r}
new_df = new_df %>%
  filter(!is.na(Male_Weekly_Income))
new_df = new_df %>%
  filter(!is.na(Female_Weekly_Income))
new_df
```

Now that the data is tidy and transformed, we can create a new column that specifies the difference between the two columns. 
```{r}
new_df$Difference = new_df$Male_Weekly_Income - new_df$Female_Weekly_Income
new_df
```
From the Difference column, we can see the difference between weekly income from males and females. The data shows males consistently make a higher weekly income than females. Additionally, I added a percentage difference in weekly income between the two genders. We can analyze the percentage amount males get paid more than females per occupation.
```{r}
new_df$Total = new_df$Male_Weekly_Income + new_df$Female_Weekly_Income
new_df$Percentage = ((new_df$Male_Weekly_Income / new_df$Female_Weekly_Income)*100)-100
new_df
```

